Figures from: Vertebrate host availability and the thermal properties of mosquito-borne parasite transmission

### 2023-12-10 manuscript-figures.Rmd
clean up code

# Vector trait TPC data
data.Vec <- read_rds("results/VecTPC_vals.rds") %>%
  pathosystem_labeller()

@@

# Uncertainty data for Topt
Topt.HPD <- read_rds("results/Topt_HPD_unc.rds") %>%
  ungroup() %>%
  pathosystem_labeller() %>%
  focalvar_labeller()

@@

```{r fig-S8_2_dR0dvar_uncertainty, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_unc} The local sensitivity of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of $\\sigma_H$. Generally, $\\mathcal{R}_0$ is most sensitive to biting frequency, the probability of surviving the extrinsic incubation period, and vector competence. Shown here are the median values of the derivative of $\\mathcal{R}_0$ with respect to each focal mosquito trait across the mosquito trait TPC hyperparameter posterior distributions. Sensitivity values are shown on a logarithmic scale with solid lines indicating positive values and dotted lines indicating negative. Temperatures are restricted to those at which the median of $\\mathcal{R}_0$ is positive for the given system. '}
KH_vals <- c(1)

temp_lims <- data.dR0dvar %>% 
  filter(focal_var == "Adult mortality rate") %>% 
  filter(!is.na(median)) %>% 
  right_join(select(data.R0, system_label, Temperature, Model, sigmaH, KH, median_R0 = median) %>%
               filter(KH %in% KH_vals & sigmaH %in% c(100, Inf))) %>%
  filter(median_R0 > 0) %>%
  group_by(system_ID) %>% 
  summarise(min_Temp = min(Temperature),
            max_Temp = max(Temperature))

val_lims <- tibble(sigmaH = c(100, Inf), 
                   min_median = c(-90, -0.05), 
                   max_median = c(5, 0.5))

dR0dk.df <- data.dR0dvar %>%
  filter(
    KH %in% KH_vals,
    sigmaH %in% c(100, Inf)
  ) %>%
  right_join(select(data.R0, system_label, Temperature, Model, sigmaH, KH, median_R0 = median) %>%
               filter(KH %in% KH_vals & sigmaH %in% c(100, Inf))) %>%
  right_join(temp_lims) %>% 
  group_by(system_ID) %>% 
  filter(Temperature > min_Temp, Temperature < max_Temp) %>% 
  ungroup() %>% 
  right_join(val_lims) %>% 
  group_by(sigmaH) %>% 
  filter(median > min_median, median < max_median) %>% 
  # filter(median_R0 > 0) %>%
  mutate(abs_median = abs(median)) %>%
  # filter(abs_median < 1e10) %>%
  mutate(sign_median = sign(median)) %>%
  filter(sign_median != 0, !is.na(sign_median)) %>%
  arrange(sign_median) %>%
  arrange(Temperature)

plot.dR0dk <- dR0dk.df %>%
  # filter(focal_var %in% c("Adult mortality rate")) %>%
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = median, color = focal_var, 
             # linetype = (as.factor(sign_median))
             )) +
  geom_line(lwd = 1) +
  scale_color_manual(
    name = "Focal trait:",
    values = rev(c(c4a("misc.okabe", 7), "black"))
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(
    name = unname(TeX("Local sensitivity of $R_0$")),
    # limits = c(-1, 1),
    # limits = 10^c(-4, 4),
    # breaks = 10^(seq(-4, 4)),
    # labels = TeX(c("$10^{-4}$", "$10^{-3}$", "$10^{-2}$", "$10^{-1}$", "$10^0$", "$10^1$", "$10^2$", "$10^3$", "$10^4$")),
    # trans = "log10",
    expand = c(0, 0)
  ) +
  scale_linetype_manual(
    name = "Sign of sensitivity:",
    values = c(1, 3),
    breaks = c(1, -1),
    labels = c("Positive", "Negative")
  ) +
  facet_rep_grid(
    cols = vars(system_label),
    rows = vars(KH, sigmaH),
    labeller = labeller(
      system_label = sens_labels,
      sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
      KH = as_labeller(appender_KH, default = label_parsed)
    ),
    scales = "free"
  ) +
  guides(linetype = guide_legend(
    title.position = "left",
    title.vjust = 0.5,
    direction = "vertical"
  )) +
  theme_minimal_hgrid(10) +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.position = "bottom",
    legend.justification = 0.5,
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 10,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

@@

>>>>

# Vector trait TPC data

# Create data frame of TPCs

data.Vec <- read_rds("results/VecTPC_vals.rds") %>%
  pathosystem_labeller()

thin_size <- 500
num_samples <- length(unique(data.Vec$sample_num))
sample_inds <- sample(1:num_samples, thin_size, replace = FALSE)

data.Vec <- filter(data.Vec, sample_num %in% sample_inds)

@@

# Uncertainty data for Topt
Topt.HPD <- read_rds("results/Topt_HPD_unc.rds") %>%
  mutate(focal_var = ifelse(focal_var == "lf", "muV", focal_var)) %>% 
  ungroup() %>%
  pathosystem_labeller() %>%
  focalvar_labeller()

@@

```{r fig-S8_2_dR0dvar_uncertainty, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_unc} The local sensitivity of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of $\\sigma_H$ and $K_H$. Generally, $\\mathcal{R}_0$ is most sensitive to biting frequency, the probability of surviving the extrinsic incubation period, and vector competence. Shown here are the median values of the derivative of $\\mathcal{R}_0$ with respect to each focal mosquito trait across the mosquito trait TPC hyperparameter posterior distributions. Sensitivity values are shown on a logarithmic scale with solid lines indicating positive values and dotted lines indicating negative. Temperatures are restricted to those at which the median of $\\mathcal{R}_0$ is positive for the given system. '}
KH_vals <- c(1, 100)

data.dR0dT <- data.dR0dvar %>%
  filter((KH %in% KH_vals & sigmaH == 10) | (KH == 100 & is.infinite(sigmaH)))

temp_lims <- select(data.R0, system_ID, system_label, Temperature, Model, sigmaH, KH, median_R0 = median) %>%
  filter(KH %in% KH_vals & sigmaH %in% c(100, Inf)) %>%
  filter(!is.na(median_R0), median_R0 > eps) %>%
  group_by(system_ID) %>% 
  summarise(min_Temp = min(Temperature),
            max_Temp = max(Temperature))

val_lims <- tibble(sigmaH = unique(data.dR0dT$sigmaH), 
                   min_median = c(-90, -0.05), 
                   max_median = c(5, 0.5))

dR0dk.df <- data.dR0dT %>%
  filter(
    KH %in% KH_vals,
    sigmaH %in% unique(data.dR0dT$sigmaH)
  ) %>%
  right_join(select(data.R0, system_ID, system_label, Temperature, Model, sigmaH, KH, median_R0 = median) %>%
               filter(KH %in% KH_vals & sigmaH %in% unique(data.dR0dT$sigmaH))) %>%
  # Restrict to relevant temperatures (V0 and R0 > 0)
  right_join(temp_lims) %>% 
  group_by(system_ID) %>% 
  filter(Temperature > min_Temp, Temperature < max_Temp) %>% 
  filter(!is.na(median)) %>% 
  ungroup() %>% 
  # Restrict plotted values (so all plots are visible)
  right_join(val_lims) %>% 
  group_by(sigmaH) %>% 
  # filter(median > min_median, median < max_median) %>% 
  ungroup() %>% 
  # filter(median_R0 > 0) %>%
  mutate(abs_median = abs(median)) %>%
  # Normalize values to better tell which is most important
  mutate(sign_median = sign(median)) %>%
  group_by(system_ID, sigmaH, KH) %>%
  mutate(norm_median = ifelse(sign_median>0,
                              sign_median * median / max(median, na.rm = TRUE),
                              sign_median * median / min(median, na.rm = TRUE))) %>%
  filter(sign_median != 0, !is.na(sign_median)) %>%
  arrange(sign_median) %>%
  arrange(Temperature)

# # R0 derivative dataframe
# R0deriv_df <- data.R0 %>% 
#   group_by(system_ID, Model, sigmaH, KH) %>% 
#   arrange(system_ID, Model, sigmaH, KH, Temperature) %>%
#   mutate(deriv = (mean - lag(mean)) / (Temperature - lag(Temperature)))

# test_plot <- R0deriv_df %>% 
#   filter(KH %in% c(1), sigmaH %in% c(100, Inf), median > 0) %>% 
#   ggplot(aes(x = Temperature, y = deriv)) +
#   geom_line() +
#   geom_hline(yintercept = 0) +
#   facet_rep_grid(
#     rows = vars(KH, sigmaH),
#     cols = vars(system_label),
#     labeller = labeller(
#       system_label = sens_labels,
#       sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
#       KH = as_labeller(appender_KH, default = label_parsed)
#     ),
#     scales = "free"
#   ) 
# test_plot

# NB: in order to produce an R0 TPC, overall derivative should go from positive to negative
plot.dR0dk <- dR0dk.df %>%
  # filter(focal_var %in% c("Overall")) %>%
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = norm_median, color = focal_var, 
             # linetype = (as.factor(sign_median))
  )) +
  geom_line(lwd = 0.5) +
  geom_hline(yintercept = 0, lwd = 0.25) +
  scale_color_manual(
    name = "Focal trait:",
    values = rev(c(c4a("misc.okabe", 7), "black"))
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(
    name = unname(TeX("Normalized local sensitivity of $\\,R_0$ to temperature")),
    # limits = c(-1, 1),
    # limits = 10^c(-4, 4),
    # breaks = 10^(seq(-4, 4)),
    # labels = TeX(c("$10^{-4}$", "$10^{-3}$", "$10^{-2}$", "$10^{-1}$", "$10^0$", "$10^1$", "$10^2$", "$10^3$", "$10^4$")),
    # trans = "log10",
    expand = c(0, 0)
  ) +
  # scale_linetype_manual(
  #   name = "Sign of sensitivity:",
  #   values = c(1, 3),
  #   breaks = c(1, -1),
  #   labels = c("Positive", "Negative")
  # ) +
  facet_rep_grid(
    rows = vars(KH, sigmaH),
    cols = vars(system_label),
    labeller = labeller(
      system_label = sens_labels,
      sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
      KH = as_labeller(appender_KH, default = label_parsed)
    ),
    scales = "free"
  ) +
  guides(linetype = guide_legend(
    title.position = "left",
    title.vjust = 0.5,
    direction = "vertical"
  )) +
  theme_minimal_hgrid(10) +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.position = "bottom",
    legend.justification = 0.5,
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 10,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

@@

```{r fig-S8_3_dR0dvar_comparison, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_comp} Comparison of the local sensitivities of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of $\\sigma_H$. Generally, $\\mathcal{R}_0$ is most sensitive to biting frequency, the probability of surviving the extrinsic incubation period, and vector competence. Shown here are the median values of the derivative of $\\mathcal{R}_0$ with respect to each focal mosquito trait across the mosquito trait TPC hyperparameter posterior distributions. Sensitivity values are shown on a logarithmic scale with solid lines indicating positive values and dotted lines indicating negative. Temperatures are restricted to those at which the median of $\\mathcal{R}_0$ is positive for the given system. '}

comp_dR0dk.df <- dR0dk.df %>%
  # filter(sigmaH == 10) %>% 
  filter(focal_var != "Overall") %>% 
  group_by(system_ID, Temperature, sigmaH, KH) %>% 
  mutate(abs_median = abs(median),
         total_sens = sum(abs_median),
         prop_sens = abs_median / total_sens) %>% 
  arrange(system_label, Temperature, KH, sigmaH, focal_var) #%>% 
# filter(focal_var %in% c("Parasite development rate", "Biting frequency"))
# # mutate(median = abs(median)) %>% 
# select(system_ID, Temperature, sigmaH, KH, focal_var, system_label, prop_sens) %>%
# pivot_wider(names_from = KH, values_from = prop_sens) %>% 
# mutate(sens_comp = ifelse(`10` == 0 & `1` == 0,
#                           0,
#                           (`10` - `1`) / ((`10` + `1`)/2)))
# mutate(sens_comp = (`10` - `1`) / `1`)

plot.comp_dR0dk <- comp_dR0dk.df %>%
  # filter(focal_var %in% c("Overall")) %>%
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = prop_sens, fill = focal_var, color = focal_var,
             # linetype = (as.factor(sign_median))
  )) +
  # geom_area() +
  geom_line(lwd = 1) +
  geom_hline(yintercept = 0, lwd = 0.25) +
  scale_color_manual(
    name = "Focal trait:",
    values = rev(c4a("misc.okabe", 7))
  ) +
  scale_fill_manual(
    name = "Focal trait:",
    values = rev(c4a("misc.okabe", 7))
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(
    name = unname(TeX("Contributions to overall local sensitivity of $\\,R_0$")),
    # limits = c(-100, 100),
    # limits = c(-0.2, 0.1),
    # limits = 10^c(-4, 4),
    # breaks = 10^(seq(-4, 4)),
    # labels = TeX(c("$10^{-4}$", "$10^{-3}$", "$10^{-2}$", "$10^{-1}$", "$10^0$", "$10^1$", "$10^2$", "$10^3$", "$10^4$")),
    # trans = "log10",
    expand = c(0, 0)
  ) +
  # scale_linetype_manual(
  #   name = "Sign of sensitivity:",
  #   values = c(1, 3),
  #   breaks = c(1, -1),
  #   labels = c("Positive", "Negative")
  # ) +
  facet_rep_grid(
    rows = vars(sigmaH, KH),
    cols = vars(system_label),
    labeller = labeller(
      system_label = sens_labels,
      sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
      KH = as_labeller(appender_KH, default = label_parsed)
    ),
    scales = "free"
  ) +
  guides(linetype = guide_legend(
    title.position = "left",
    title.vjust = 0.5,
    direction = "vertical"
  )) +
  theme_minimal_hgrid(10) +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.position = "bottom",
    legend.justification = 0.5,
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 10,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

plot.comp_dR0dk
```

```{r fig-S8_4_dR0dvar_compzoom, dev = c('pdf','svg', 'png'), message=FALSE, out.width='100%', fig.width=9, fig.height=6, fig.align='center', fig.cap='\\label{fig:dR0dvar_comp} Comparison of the local sensitivities of $\\mathcal{R}_0$ to mosquito traits across temperatures for two values of $\\sigma_H$. Generally, $\\mathcal{R}_0$ is most sensitive to biting frequency, the probability of surviving the extrinsic incubation period, and vector competence. Shown here are the median values of the derivative of $\\mathcal{R}_0$ with respect to each focal mosquito trait across the mosquito trait TPC hyperparameter posterior distributions. Sensitivity values are shown on a logarithmic scale with solid lines indicating positive values and dotted lines indicating negative. Temperatures are restricted to those at which the median of $\\mathcal{R}_0$ is positive for the given system. '}

compzoom_dR0dk.df <- comp_dR0dk.df %>%
  filter(sigmaH == 10) %>% 
  select(system_ID, Temperature, sigmaH, KH, focal_var, system_label, prop_sens) %>%
  pivot_wider(names_from = KH, values_from = prop_sens) %>%
  mutate(abs_comp = abs(`1`) - abs(`100`))

plot.compzoom_dR0dk <- compzoom_dR0dk.df %>% 
  # filter(focal_var %in% c("Parasite development rate", "Adult mortality rate", "Biting frequency")) %>%
  # Make a stacked plot, colored by parameter
  ggplot(aes(x = Temperature, y = abs_comp, fill = focal_var, color = focal_var,
             # linetype = (as.factor(sign_median))
  )) +
  # geom_area() +
  # geom_smooth(se = FALSE) +
  geom_line(lwd = 1) +
  geom_hline(yintercept = 0, lwd = 0.25) +
  scale_color_manual(
    name = "Focal trait:",
    values = rev(c4a("misc.okabe", 7))
  ) +
  scale_fill_manual(
    name = "Focal trait:",
    values = rev(c4a("misc.okabe", 7))
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(
    name = unname(TeX("Change in contributions to overall local sensitivity of $\\,R_0$ \n  from $\\,K_H = 1$ to $\\,K_H = 10$ ")),
    # limits = c(-100, 100),
    # limits = c(-0.2, 0.1),
    # limits = 10^c(-4, 4),
    # breaks = 10^(seq(-4, 4)),
    # labels = TeX(c("$10^{-4}$", "$10^{-3}$", "$10^{-2}$", "$10^{-1}$", "$10^0$", "$10^1$", "$10^2$", "$10^3$", "$10^4$")),
    # trans = "log10",
    expand = c(0, 0)
  ) +
  # scale_linetype_manual(
  #   name = "Sign of sensitivity:",
  #   values = c(1, 3),
  #   breaks = c(1, -1),
  #   labels = c("Positive", "Negative")
  # ) +
  facet_rep_grid(
    rows = vars(sigmaH),
    cols = vars(system_label),
    labeller = labeller(
      system_label = sens_labels,
      # sigmaH = as_labeller(appender_sigmaH, default = label_parsed),
      # KH = as_labeller(appender_KH, default = label_parsed)
    ),
    scales = "free"
  ) +
  guides(linetype = guide_legend(
    title.position = "left",
    title.vjust = 0.5,
    direction = "vertical"
  )) +
  theme_minimal_hgrid(10) +
  theme(
    legend.margin = margin(0, 0, 0, 0),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.position = "bottom",
    legend.justification = 0.5,
    axis.title.x = element_text(hjust = 0.5),
    strip.text.x = ggtext::element_markdown(
      size = 10,
      hjust = 0,
      padding = margin(0, 0, 0, 0),
      margin = margin(1, 1, 1, 1)
    )
  )

plot.compzoom_dR0dk
```
